Context Inconsistency Management \\in Pervasive Systems

### 2011-04-12 paper/paper.tex
extend results section

\subsection{Case Study}
Suppose we want to build a context-aware application that has a certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with its normal operation. In the fuzzy ontology approach these unpredictable events can be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without the need for conflict resolution. A simplified hidden Markov model for this ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying teachers activities and the probabilities of different locations in the presence of this higher level context.}
  \label{fig:fuzzy}
\end{figure}
As the numbers in this model indicate, the most probable location of a teacher while lecturing is at the lectern (L). In the aisle (A) is less probable and at the office (O) is very unlikely. For the activity of performing other work (denoted as Working in the model), the probabilities are the other way around. This fuzzy ontology model can now be used to calculate the relative probability of each higher-level context (c.q. Lecturing and Working) based a set of low level contexts observed within a certain time window. A possible set of observations could be [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$]: the teacher was observed three times at the lectern and twice in the aisle in this time window. The probability of the Lecturing activity can now be calculated by multiplying the probabilities of the observations in the presence of this higher level context\footnote{We multiply the results with a constant $c$ in order to obtain a valid probability value in the end}: $pLecturing = .80 * .80 * .80 * .15 * .15 * c_1 = .01152 * c_1 = 0.9998$. The probability for Working can be calculated in a similar way: $pWorking = .05 * .05 * .05 * .15 * .15 * c_1 = 0.000002813 * c_1 = 0.0002$. The probabilities indicate that the teacher is still lecturing, even though the teacher being observed in the aisle conflicts with being observed at the lectern. Another example of observations could be [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$]: the teacher is observed at the lectern once, twice in the aisle and twice at the office. The probabilities for Lecturing and Working are calculated as follows: $pLecturing = .80 * .15 * .15 * .05 * .05 * c_2 = 0.000045 * c_2 = 0.059$ and $pWorking = .05 * .15 * .15 * .80 * .80 * c_2 = 0.00072 * c_2 = 0.941$. The probabilities now indicate that the teacher is doing work other than lecturing, even though the observation of the teacher at the lectern is definitely conflicting with the observations at the office.
\subsection{Results}
The case study shows that a fuzzy ontology allows us to always define the most probable higher-level context, regardless of any inconsistencies in the observed context set and without the need for conflict detection and resolution.

Another observation we can make is that the hidden Markov model for a fuzzy ontology has some tree-like structure. This structure allows for very efficient incremental partial probability calculation (PPC).

The probability of a higher-level context is calculated by simply multiplying the probabilities of the observed contexts. When a new context is added to the observation window, the intermediate calculation result stored at a node (the root in our case study) can simply be multiplied by the probability of the new observation in order to obtain the updated probability for a higher-level context. When a context disappears from the observation window, the intermediate result can simply be divided by the context's probability to obtain an updated probability for a higher-level context.

Note that it is possible to extend the simple fuzzy ontology from the use case by specifying start and transition probabilities for the higher-level contexts, just as hidden Markov models do, and other information such as the expected lecture time window (fuzzified of course~\cite{Buckley1995245}).

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change frequently. Partial constraint checking was 15 times as efficient as the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context aware application that has to handle frequently generated contexts.

After that we shed light on ways to resolve the detected conflicts and discussed a context inconsistency resolution algorithm which estimates the reliability of conflicting contexts and only retains the most reliable context from every conflict set. Using this algorithm allows for timely conflict resolution at runtime without the need for human participation. The algorithm is a valuable addition to the research on conflict resolution in pervasive computing, but when applying it one has to verify the aptness of the heuristic used for context reliability estimations in the domain at hand. In particular the assumption that contexts with the highest relative frequency are most reliable has to apply in the domain.

Finally the use of fuzzy ontologies for modeling the environment of context-aware applications seems promising. By incorporating possible inconsistencies in the model the need for context inconsistency management vanishes, vastly reducing the complexity of the used middleware. Fuzzy ontologies always yield the most probable higher-level contexts, regardless of the inconsistencies in the observed context set. In addition the calculation of the probabilities in a fuzzy ontology has great potential for optimization by using incremental partial probability calculation. 
When using fuzzy ontologies one however has to take into account the required training of the system for obtaining well defined probabilities in the context model.

>>>>

\subsection{Case Study}
Suppose we want to build a context-aware application that has a certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with its normal operation. In the fuzzy ontology approach these unpredictable events can be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without the need for conflict resolution. A simplified hidden Markov model for this ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying a teacher's activities and the probabilities of different locations in the presence of these higher-level context.}
  \label{fig:fuzzy}
\end{figure}
As the numbers in this model indicate, the most probable location of a teacher while lecturing is at the lectern (L). In the aisle (A) is less probable and at the office (O) is very unlikely. For the activity of performing other work (denoted as Working in the model), the probabilities are the other way around. This fuzzy ontology model can now be used to calculate the relative probability of each higher-level context (c.q. Lecturing and Working) based a set of low level contexts observed within a certain time window. A possible set of observations could be [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$]: the teacher was observed three times at the lectern and twice in the aisle in this time window. The probability of the Lecturing activity can now be calculated by multiplying the probabilities of the observations in the presence of this higher level context\footnote{We multiply the results with a constant $c$ to obtain a valid probability value in the end}: $pLecturing = .80 * .80 * .80 * .15 * .15 * c_1 = .01152 * c_1 = 0.9998$. The probability for Working can be calculated in a similar way: $pWorking = .05 * .05 * .05 * .15 * .15 * c_1 = 0.000002813 * c_1 = 0.0002$. The probabilities indicate that the teacher is still lecturing, even though the teacher being observed in the aisle conflicts with being observed at the lectern. Another example of observations could be [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$]: the teacher is observed at the lectern once, twice in the aisle and twice at the office. The probabilities for Lecturing and Working are calculated as follows: $pLecturing = .80 * .15 * .15 * .05 * .05 * c_2 = 0.000045 * c_2 = 0.059$ and $pWorking = .05 * .15 * .15 * .80 * .80 * c_2 = 0.00072 * c_2 = 0.941$. The probabilities now indicate that the teacher is doing work other than lecturing, even though the observation of the teacher at the lectern is definitely conflicting with the observations at the office.
\subsection{Results}
The case study shows that a fuzzy ontology allows us to always define the most probable higher-level context given the observations, regardless of any inconsistencies in the observed context set and without the need for conflict detection and resolution.

Another observation we can do is that the hidden Markov model for a fuzzy ontology has a tree-like structure. This structure allows for very efficient incremental Partial Probability Calculation (PPC). The PCC principle works as follows: the probability of a high-level context is calculated by multiplying the probabilities of its children contexts. When a new context is added to or removed from the observation set and thus from the tree, only the probabilities of the observation's ancestors in the tree up to the root have to be recalculated. In case of adding a context, the probability of an ancestor in the tree can be recalculated by simply multiplying its previous probability with the probability of the new or updated child. In case of removing a context, the probability is divided by the probability of the new or updated child. Because the recalculation only takes place in part of the tree and allows us to preserver previous calculations for later use, it is partial. And because this recalculation only happens in trees that contain the added or removed context as a leaf, the calculation is incremental. We illustrate the PPC process using an extension of the hidden Markov model from the case study (See Fig.~\ref{fig:fuzzy2}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy2}
  \caption{Fuzzy ontology specifying the Lecturing activity and the probabilities of different contexts in the presence of this higher-level context.}
  \label{fig:fuzzy2}
\end{figure}
In this figure the usesSlides context is added to the observation set at some point. This change in the context set requires us to recalculate the probabilities of the higher-level contexts. For this we use PPC: we recalculate only the values of the ancestors if the usesSlides context (c.q. isPresenting and isLecturing). The probability of isPresenting is $.85 * .90$, because the usesSlides context is its only child. The probability of isLecturing has to be recalculated. This can be done by simply multiplying its previous probability with the probability of the new child isPresenting: $isLecturing = 0.0004104 * .90 * .85 = 0.000313956$\footnote{Note that this value has to be multiplied with a constant in order to get the real probability value}.

The time complexity of the PPC principle is related to the height of a hidden Markov tree, because recalculations only take place on the path from a leaf to the root. Since this height remains constant at runtime, the time complexity of adding or removing a context is $O(1)$.

It is possible to extend the hidden Markov model used in the case study by specifying start and transition probabilities for the higher-level contexts. With this extension the probability of a higher-level context does not solely depend on the contexts in the observation set, but also on the previous most probable higher-level context.

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change frequently. Partial constraint checking was 15 times more efficient than the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context aware application that has to handle frequently generated contexts.

After that we shed light on ways to resolve the detected conflicts and discussed a context inconsistency resolution algorithm which estimates the reliability of conflicting contexts and only retains the most reliable context from every conflict set. This algorithm allows for timely conflict resolution at runtime without the need for human participation. The algorithm is a valuable addition to the research on conflict resolution in pervasive computing, but when applying it one has to verify the aptness of the heuristic used for context reliability estimations in the domain at hand. In particular the assumption that contexts with the highest relative frequency are most reliable has to apply in the domain.

Finally the use of fuzzy ontologies for modeling the environment of context-aware applications seems promising. By incorporating possible inconsistencies in the model, the need for context inconsistency management vanishes, vastly reducing the complexity of the used middleware. Fuzzy ontologies always yield the most probable higher-level contexts, regardless of the inconsistencies in the observed context set. In addition the calculation of the probabilities in a fuzzy ontology has great potential for optimization by using incremental Partial Probability Calculation. 
When using fuzzy ontologies one however has to take into account the required training of the system for obtaining well defined probabilities in the context model.
